\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-05-29 doc/incertae_sedis.tex
Spacing.
Added some drafty discussion text.

\begin{figure}
%\subfloat[Input tree \#1]{\includegraphics[scale=0.5]{Figures/is-within-is/tree1\lyxdot tre}}
\subfloat[maximally scored]{
    \begin{tabular}{|r|p{2em}|p{2em}|}
\hline
& \multicolumn{2}{|c|}{Character} \\
taxa & $V$ & $WX$  \\
\hline
$U$  & 0 & 0  \\
$V1$ & 1 & ?  \\
$V2$ & 1 & ?  \\
$W$  & 0 & 1  \\
$X$  & 0 & 1 \\
\hline
\end{tabular}
}
\subfloat[less informative]{
    \begin{tabular}{|r|p{2em}|p{2em}|}
\hline
& \multicolumn{2}{|c|}{Character} \\
taxa & $V$ & $WX$  \\
\hline
$U$  & 0 & 0 \\
$V1$ & 1 & ? \\
$V2$ & 1 & ? \\
$W$  & ? & 1 \\
$X$  & 0 & 1 \\
\hline
\end{tabular}
}
\subfloat[from revised taxonomy]{
    \begin{tabular}{|r|p{2em}|p{2em}|}
\hline
& \multicolumn{2}{|c|}{Character} \\
taxa & $V$ & $WX$  \\
\hline
$U$  & 0 & 0 \\
$V1$ & 1 & 1 \\
$V2$ & 1 & 1 \\
$W$  & ? & 1 \\
$X$  & 0 & 1 \\
\hline
\end{tabular}
}
\caption{\label{fig:Placing-one-incertae-MRP}
Panels (a) and (b) show two possible encodings that would support the $V$ and $WX$ branches and \emph{incertae sedis}
annotation shown in the taxonomy from in Figure \ref{fig:Placing-one-incertae}a.
The representations only vary in the scoring of leaf $W$ for character $V$. Panel (c) shows the information encoded by the tree in \ref{fig:Placing-one-incertae}c. }
\end{figure}

@@

\begin{figure}
\subfloat[$\tau_{1}$]{\includegraphics[scale=0.5]{Figures/decomp/tree1}
}
\subfloat[$\tau_{2}$]{\includegraphics[scale=0.5]{Figures/decomp/tree2}
}
\subfloat[$\tau_{3}$]{\includegraphics[scale=0.5]{Figures/decomp/tree3}
}
\subfloat[$\tau_{4}$]{\includegraphics[scale=0.5]{Figures/decomp/tree4}
}

\subfloat[Taxonomy $\protect\taxonomy$]{\includegraphics[scale=0.5]{Figures/decomp/tax}
}
\subfloat[$Y$ and $Z$ decomposition]{\includegraphics[scale=0.5]{Figures/decomp/decompRep}
}
\subfloat[Pruned $Y$ and $Z$ decomposition]{\includegraphics[scale=0.5]{Figures/decomp/decompPruned}
}

@@

\textbf{BDR:}\textbf{\emph{ }}\emph{we might get this behavior in a
nice an automatic way if we create a single fake leaf for each
monotypic taxonomy node that holds the node's leaf label.}

@@

>>>>

\begin{figure}
%\subfloat[Input tree \#1]{\includegraphics[scale=0.5]{Figures/is-within-is/tree1\lyxdot tre}}
\subfloat[maximally scored]{
    \begin{tabular}{|r|p{2em}|p{2em}|}
\hline
& \multicolumn{2}{|c|}{Character} \\
taxa & $V$ & $WX$  \\
\hline
$U$  & 0 & 0  \\
$V1$ & 1 & ?  \\
$V2$ & 1 & ?  \\
$W$  & 0 & 1  \\
$X$  & 0 & 1 \\
\hline
\end{tabular}
}\hfill{}
\subfloat[less informative]{
    \begin{tabular}{|r|p{2em}|p{2em}|}
\hline
& \multicolumn{2}{|c|}{Character} \\
taxa & $V$ & $WX$  \\
\hline
$U$  & 0 & 0 \\
$V1$ & 1 & ? \\
$V2$ & 1 & ? \\
$W$  & ? & 1 \\
$X$  & 0 & 1 \\
\hline
\end{tabular}
}\hfill()
\subfloat[from revised taxonomy]{
    \begin{tabular}{|r|p{2em}|p{2em}|}
\hline
& \multicolumn{2}{|c|}{Character} \\
taxa & $V$ & $WX$  \\
\hline
$U$  & 0 & 0 \\
$V1$ & 1 & 1 \\
$V2$ & 1 & 1 \\
$W$  & ? & 1 \\
$X$  & 0 & 1 \\
\hline
\end{tabular}
}
\caption{\label{fig:Placing-one-incertae-MRP}
Panels (a) and (b) show two possible encodings that would support the $V$ and $WX$ branches and \emph{incertae sedis}
annotation shown in the taxonomy from in Figure \ref{fig:Placing-one-incertae}a.
The representations only vary in the scoring of leaf $W$ for character $V$. Panel (c) shows the information encoded by the tree in \ref{fig:Placing-one-incertae}c. }
\end{figure}

@@

\begin{figure}
\subfloat[$\tau_{1}$]{\includegraphics[scale=0.5]{Figures/decomp/tree1}
}\hfill{}
\subfloat[$\tau_{2}$]{\includegraphics[scale=0.5]{Figures/decomp/tree2}
}\hfill{}
\subfloat[$\tau_{3}$]{\includegraphics[scale=0.5]{Figures/decomp/tree3}
}\hfill{}
\subfloat[$\tau_{4}$]{\includegraphics[scale=0.5]{Figures/decomp/tree4}
}

\subfloat[Taxonomy $\protect\taxonomy$]{\includegraphics[scale=0.5]{Figures/decomp/tax}
}\hfill{}
\subfloat[$Y$ and $Z$ decomposition]{\includegraphics[scale=0.5]{Figures/decomp/decompRep}
}\hfill{}
\subfloat[Pruned $Y$ and $Z$ decomposition]{\includegraphics[scale=0.5]{Figures/decomp/decompPruned}
}

@@

\bdrcomment{we might get this behavior in a
nice an automatic way if we create a single fake leaf for each
monotypic taxonomy node that holds the node's leaf label.}

@@

\section{Discussion}

In this paper we've described a method for handling \emph{incertae sedis} taxa
when constructing taxonomic supertrees.  Our method incorporates the uncertain
placement information from such taxa when constructing the supertree.  It also
handles higher taxon names and allows input trees to revise the taxonomy by
determining where \emph{incertae sedis} taxa should be placed.

However, our method leaves a number of problems for future work.  One of these
problems is that we consider taxa with an \emph{incertae sedis} label to 
about which siblings in attach within.  However, in practice,
the taxonomist may only consider it plausible for the taxon to intrude only within specific
siblings.  Our method also assumes that an \emph{incertae sedis} label on a taxon
indicates that a taxon may attach at any depth within its siblings.  In practice, an
\emph{incertae sedis} genus with sister families (for example) is unlikely to attach
so deeply within a sister family that it becomes nested within another genus.

In order to consider more precise attachment ranges for \emph{incertae sedis} taxa, we
would have to obtain specifications for attachment ranges that are machine readable and
robust.  Obtaining these specifications would be difficult or impossible in many cases.
However, in cases that result from taxonomy merging (below) it would be possible to obtain
them.  This necessitates the creation of a format \emph{incertae sedis} taxa that is more
expressive that a simple label.
